### 2023-07-05 chapters/interpretability/relaxed-cnf/experiment.tex
addressing comments

\begin{table}
	\caption[Accuracy, rule-size, and training time of rule-based classifiers]{Comparisons of test accuracy, rule-size and training time among different rule-based classifiers. Every cell in the last four columns contains the test accuracy in percentage (top value), rule size  (middle value), and  training time in seconds (bottom value). In the experiments, {\crr} shows higher test accuracy than IMLI and generates sparser rules than RIPPER. Number in bold denotes the best result, such as maximum test accuracy, minimum rule-size, and minimum training time among competitive classifiers.}
	\label{interpretability_crr_tab:rule_based_classifiers}

@@

%%\vspace{-9pt}
\subsection{Experimental Results}
In the following, we first discuss empirical results of rule-based classifiers, then extend analysis to non-rule-based classifiers, and finally discuss the effect of different choices of hyper-parameters. 
%%\vspace{-3pt}
\subsubsection{Performance Evaluation of {\crr} with Rule-based Classifiers}
We conduct an assessment of performance  using five-fold nested cross-validation as in~\cite{DGW2018}  and report the median of test accuracy, rule-size and training time of all rule-based classifiers in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. 
Specifically, we show the dataset,  the number of samples and the number of discretized features in the first three columns in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. Inside each cell of column   four to $    11 $,  we present the   test accuracy (top value), rule-size (middle value) and training time (bottom value) of each classifier for each dataset.

We first compare relaxed-CNF rules generated by {\crr} with CNF rules generated by IMLI. In {Table}~\ref{interpretability_crr_tab:rule_based_classifiers}, relaxed-CNF rules exhibit higher prediction accuracy than CNF rules in the majority of the datasets, showing the effectiveness of using a more expressive representation of classification rules in capturing the decision boundary. In addition, the generated relaxed-CNF rules are comparatively smaller than CNF rules in terms of rule-size in most of the datasets. Therefore, relaxed-CNF rules improve upon CNF rules in terms of both prediction accuracy and rule size in the majority of the datasets. In this context, {\crr} provides a trade-off between accuracy and rule-size depending on the choice of hyper-parameters and the experimental results are discussed later in Chapter~\ref{interpretability_crr_sec:model_parameters}.  We then compare relaxed-CNF rules with DNF rules generated by BRS and find that relaxed-CNF rules outperform DNF rules with respect to prediction accuracy in several datasets. At this point,   BRS fails to scale on larger datasets as shown in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. We finally compare relaxed-CNF rules with decision lists generated by RIPPER.  In the experiments, relaxed-CNF rules achieve comparable prediction accuracy with decision lists in most of the datasets. In contrast,   RIPPER generates very large decision lists compared to relaxed-CNF rules in the majority of the datasets, more precisely in large datasets.  To summarize the performance of {\crr} among different rule-based classifiers,  {\crr} can generate smaller relaxed-CNF rules with better accuracy in numbers of the cases with a couple of exceptions.

@@

	\caption[Effect of data-fidelity $ \lambda $ in {\crr}]{Effect of data-fidelity $ \lambda $ on test accuracy, rule-size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_lambda}
\end{figure}

@@

	\caption[Effect of the number of clause $ k $ in {\crr}]{Effect of the number of clause $ k $ on test accuracy, rule-size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_clause}
\end{figure}

@@

	\caption[Effect of mini-batch size in {\crr}]{Effect of mini-batch size on test accuracy, rule-size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_minibatch}
\end{figure}

@@

	\caption[Effect of the number of iterations in {\crr}]{Effect of the number of iterations on test accuracy, rule-size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_iteration}
\end{figure}

@@

	\paragraph{Varying the number of clauses ($ k $): }
	In Figure~\ref{interpretability_crr_fig:result_clause}, as we increase  $ k $, {\crr} allows the generated rules to capture the variance in the given dataset more effectively, that results in higher  accuracy in most of the datasets. The  rule-size also increases as we learn more clauses. In addition, the training time  increases, that can be reasoned by the fact that the number of constraints in the MILP formulation is linear with $ k $. Thereby, the number of clauses $ k $ provides a control over the accuracy of the generated rule and training time.

@@

	%	%%\vspace{-3pt}
	\paragraph{Varying the number of iterations ($ \tau $):}   
	In Figure~\ref{interpretability_crr_fig:result_iteration}, as we allow more iterations in the learning process, we find an increase in accuracy in most datasets. The training time also increases with $ \tau $ because {\crr} is required to solve in total $ \tau $ queries. We also observe an increase in rule-size in most datasets. The reason is that the objective function in the incremental mini-batch approach in Eq.~\ref{interpretability_crr_eq:ilp_inc} does not put a restriction on the size of the rules, rather on the change of rules in consecutive iterations.  In addition, the learned values of the thresholds $ \eta_c$ and $ \eta_l $  in one iteration are not carried to the MILP query in the next iteration, that may cause an increase of rule-size.

>>>>

\begin{table}
	\caption[Accuracy, rule-size, and training time of rule-based classifiers]{Comparisons of test accuracy, rule size and training time among different rule-based classifiers. Every cell in the last four columns contains the test accuracy in percentage (top value), rule size  (middle value), and  training time in seconds (bottom value). In the experiments, {\crr} shows higher test accuracy than IMLI and generates sparser rules than RIPPER. Number in bold denotes the best result, such as maximum test accuracy, minimum rule size, and minimum training time among competitive classifiers.}
	\label{interpretability_crr_tab:rule_based_classifiers}

@@

%%\vspace{-9pt}
\subsection{Experimental Results}
In the following, we first discuss empirical results of rule-based classifiers, then extend analysis to non-rule-based classifiers, and finally discuss the effect of different choices of hyper-parameters. 
%%\vspace{-3pt}
\subsubsection{Performance Evaluation of {\crr} with Rule-based Classifiers}
We conduct an assessment of performance  using five-fold nested cross-validation as in~\cite{DGW2018}  and report the median of test accuracy, rule size and training time of all rule-based classifiers in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. 
Specifically, we show the dataset,  the number of samples and the number of discretized features in the first three columns in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. Inside each cell of column   four to $    11 $,  we present the   test accuracy (top value), rule size (middle value) and training time (bottom value) of each classifier for each dataset.

We first compare relaxed-CNF rules generated by {\crr} with CNF rules generated by IMLI. In {Table}~\ref{interpretability_crr_tab:rule_based_classifiers}, relaxed-CNF rules exhibit higher prediction accuracy than CNF rules in the majority of the datasets, showing the effectiveness of using a more expressive representation of classification rules in capturing the decision boundary. In addition, the generated relaxed-CNF rules are comparatively smaller than CNF rules in terms of rule size in most of the datasets. Therefore, relaxed-CNF rules improve upon CNF rules in terms of both prediction accuracy and rule size in the majority of the datasets. In this context, {\crr} provides a trade-off between accuracy and rule size depending on the choice of hyper-parameters and the experimental results are discussed later in Chapter~\ref{interpretability_crr_sec:model_parameters}.  We then compare relaxed-CNF rules with DNF rules generated by BRS and find that relaxed-CNF rules outperform DNF rules with respect to prediction accuracy in several datasets. At this point,   BRS fails to scale on larger datasets as shown in Table~\ref{interpretability_crr_tab:rule_based_classifiers}. We finally compare relaxed-CNF rules with decision lists generated by RIPPER.  In the experiments, relaxed-CNF rules achieve comparable prediction accuracy with decision lists in most of the datasets. In contrast,   RIPPER generates very large decision lists compared to relaxed-CNF rules in the majority of the datasets, more precisely in large datasets.  To summarize the performance of {\crr} among different rule-based classifiers,  {\crr} can generate smaller relaxed-CNF rules with better accuracy in numbers of the cases with a couple of exceptions.

@@

	\caption[Effect of data-fidelity $ \lambda $ in {\crr}]{Effect of data-fidelity $ \lambda $ on test accuracy, rule size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_lambda}
\end{figure}

@@

	\caption[Effect of the number of clause $ k $ in {\crr}]{Effect of the number of clause $ k $ on test accuracy, rule size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_clause}
\end{figure}

@@

	\caption[Effect of mini-batch size in {\crr}]{Effect of mini-batch size on test accuracy, rule size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_minibatch}
\end{figure}

@@

	\caption[Effect of the number of iterations in {\crr}]{Effect of the number of iterations on test accuracy, rule size, and training time in {\crr}. } 
	\label{interpretability_crr_fig:result_iteration}
\end{figure}

@@

	\paragraph{Varying the number of clauses ($ k $): }
	In Figure~\ref{interpretability_crr_fig:result_clause}, as we increase  $ k $, {\crr} allows the generated rules to capture the variance in the given dataset more effectively, that results in higher  accuracy in most of the datasets. The  rule size also increases as we learn more clauses. In addition, the training time  increases, that can be reasoned by the fact that the number of constraints in the MILP formulation is linear with $ k $. Thereby, the number of clauses $ k $ provides a control over the accuracy of the generated rule and training time.

@@

	%	%%\vspace{-3pt}
	\paragraph{Varying the number of iterations ($ \tau $):}   
	In Figure~\ref{interpretability_crr_fig:result_iteration}, as we allow more iterations in the learning process, we find an increase in accuracy in most datasets. The training time also increases with $ \tau $ because {\crr} is required to solve in total $ \tau $ queries. We also observe an increase in rule size in most datasets. The reason is that the objective function in the incremental mini-batch approach in Eq.~\ref{interpretability_crr_eq:ilp_inc} does not put a restriction on the size of the rules, rather on the change of rules in consecutive iterations.  In addition, the learned values of the thresholds $ \eta_c$ and $ \eta_l $  in one iteration are not carried to the MILP query in the next iteration, that may cause an increase of rule size.

### 2023-07-05 chapters/interpretability/relaxed-cnf/introduction.tex
addressing comments

 {Relaxed-CNF rules are  more flexible than pure CNF rules, and they can accurately fit more complex classification boundaries.} For example, relaxed-CNF clauses  allow  a compact encoding of the majority function\footnote{The majority function is a Boolean function that evaluates to $ 0 $ when half or more arguments are false, and $ 1 $ otherwise (\url{https://en.wikipedia.org/wiki/Majority_function}).} in Boolean logic, which would require exponentially many clauses in CNF, showing the exponential gap in the succinctness of the two representations.  In addition, relaxed-CNF and CNF rules have the same functional form where a user has to compute the sum of true literals/clauses  and then compare the sum  to  different thresholds {(as in the example in Figure~\ref{interpretability_crr_exmpl:wdbc})}. From the computational perspective, the structural flexibility of relaxed-CNF compared to CNF/DNF makes it harder to learn. Therefore, in this chapter, we ask the following research question: \textit{can we design a combinatorial framework to efficiently learn relaxed-CNF rules?}

>>>>

 {Relaxed-CNF rules are  more flexible than pure CNF rules, and they can accurately fit more complex classification boundaries.} For example, relaxed-CNF clauses  allow  a compact encoding of the majority function\footnote{The majority function is a Boolean function that evaluates to $ 0 $ when half or more arguments are false, and $ 1 $ otherwise~\cite{peterson1972error}.} in Boolean logic, which would require exponentially many clauses in CNF, showing the exponential gap in the succinctness of the two representations.  In addition, relaxed-CNF and CNF rules have the same functional form where a user has to compute the sum of true literals/clauses  and then compare the sum  to  different thresholds {(as in the example in Figure~\ref{interpretability_crr_exmpl:wdbc})}. From the computational perspective, the structural flexibility of relaxed-CNF compared to CNF/DNF makes it harder to learn. Therefore, in this chapter, we ask the following research question: \textit{can we design a combinatorial framework to efficiently learn relaxed-CNF rules?}
